---
title: "Time Series Analysis in R"
date: 09.07.2019
author: "Andy Herzberg"
fig_width: 3 
fig_height: 2
output:
  html_document:
    df_print: paged
---

<a id="top"></a>

### Table of contents
* [Exploratory Data Analysis](#exploratory-data-analysis) <br>
* [Decomposition of the data](#decomposition-of-the-data) <br>
* [Autocorrelations and choosing model order](#autocorrelations)
* [Fitting an ARIMA model](#fitting-an-arima-model) <br>
* [Futher analysis and testing of our models](#further-analysis) <br>
* [Sources](#sources) <br>

<a id="exploratory-data-analysis"></a>

### Exploratory data analysis 

<a href="#top">back to top</a>

At first we need to install and load all required libraries.
```{r warning=FALSE, message=FALSE}
# install.packages("rio")
library("ggplot2")
library("forecast")
library("tseries")
library("tidyverse")
#library("rio")
library("xts")
library("dplyr")
library("readxl") #in case we need to load an Excel file
```

The data for this analysis of the Air Quality Index can be downloaded here: https://www.epa.gov/outdoor-air-quality-data/download-daily-data. 

```{r}
#Load the data
data1 <- read.csv("ad_viz_plotval_data.csv")
head(data1, 3)
```

The Date column needs to be converted to date type.

```{r}
data1$DATES <- as.Date(data1$Date, "%m/%d/%Y")
data1$Month <- format(data1$DATES,"%m")
tail(data1, 3)
```

Do some cleaning in order to keep only the relevant data bits. 

```{r}
data1$SO2 <- data1$Daily.Max.1.hour.SO2.Concentration
data1 <- select(data1, DATES, SO2, Month)
head(data1, 3)
```
Inspect the SO2 column. 

```{r}
summary(data1$SO2)
```

Draw a histogram f the SO2 data.

```{r}
ggplot(data1, aes(x=SO2)) + 
  geom_histogram(binwidth = 1, colour = "red", fill = "grey")

```


Plot the SO2 field over time - shows high variance in this case. 

```{r}
ggplot(data1, aes(DATES, SO2)) + 
  geom_line() +
  scale_x_date('Month') +
  ylab("Maximal Daily SO2 Concentration (1 hour)") +
  xlab("")
```

PLot month over month in order to see the ranges and outliers.

```{r}
ggplot(data1, aes(DATES, SO2)) + 
  geom_point(color = "red") + 
  facet_wrap(~ Month) + 
  scale_x_date('Month') + 
  ylab("Maximal Daily SO2 Concentration (1 hour)") + 
  xlab("")
```

Create a time series object based on S02 to pass to tsclean(). 

```{r}
count_TSObject <- ts(data1[,c('SO2')])
data1$clean_count = tsclean(count_TSObject)
```

Graph clean data - ignore type ts scale warning. Plot original data (blue) by the side for caomparison



















.

```{r}
 ggplot() +
  geom_line(data=data1, aes(x=DATES, y=SO2, color="Original data")) +
  geom_line(data=data1, aes(x=DATES, y=clean_count, color="Cleaned data")) +
  theme(legend.position="bottom")
```

Get weekly and monthly moving averages (MA) and compare to cleaned daily datawhich still has a lot variance and volatility in it.
```{r warning=FALSE}

data1$cnt_ma <- ma(data1$clean_count, order=7) # using the clean count with no outliers
data1$cnt_ma30 <- ma(data1$clean_count, order=30)
ggplot() +
  geom_line(data=data1, aes(x=DATES, y=clean_count, color ="Original data")) +
  geom_line(data=data1, aes(x=DATES, y=cnt_ma, color ="Weekly moving average")) +
  geom_line(data=data1, aes(x=DATES, y=cnt_ma30, color ="Monthly moving average")) +
  ylab("SO2 PPM Count") +
  theme(legend.position="bottom")
```

<a id="decomposition-of-the-data"></a>

## Decomposition of the data

<a href="#top">back to top</a>

Take seasonality, trend and cycle into account
Calculate seasonal component with stl().

```{r}
count_ma <- ts(na.omit(data1$cnt_ma), frequency=30)
decomp = stl(count_ma, s.window = "periodic")
deseasonal_cnt = seasadj(decomp)
plot(decomp)
```

Test for stationarity - first visual check for stationarity variance
2nd - Augmented Dickey-Fuller test.

```{r}
adf.test(count_ma, alternative="stationary")
```

<a id="autocorrelations"></a>

## Autocorrelations and choosing model order

<a href="#top">back to top</a>

ACF plots display correlations between a series and its lags.

```{r}
acf(count_ma, main="")
```

PACF plots display correlation between a series and its lags that explained by previous lags.

```{r}
pacf(count_ma, main="")
```

Difference of 1 is sufficient

```{r}
count_d1 = diff(deseasonal_cnt, differences = 1) 
plot(count_d1)
```

```{r}
adf.test(count_d1, alternative="stationary")
```

Look for spikes at specific lag points of the differenced series (using count_d1 from above)

```{r}
acf(count_d1, main = "ACF for differenced series")
```

```{r warning=FALSE}
pacf(count_d1, main = "PACF for differenced series")
```

<a id="fitting-an-arima-model"></a>

## Fitting an ARIMA model

<a href="#top">back to top</a>

Get auto fit p, d, q values for AR(p) I(d) MA(q)

```{r}
auto.arima(deseasonal_cnt, seasonal=FALSE)
```

Evaluate and iterate: Does the model make sense?

```{r}
fit <- auto.arima(deseasonal_cnt, seasonal=FALSE) # same as above but fititng a model
tsdisplay(residuals(fit), lag.max = 45, main='(1,1,1) Model ')
```

Graph shows series lags at 7, so modify model for p or q = 7.

```{r}
fit2 <- arima(deseasonal_cnt, order=c(1,1,7))
tsdisplay(residuals(fit2), lag.max=20, main="Seasonal Model Residuals")
```

Forecast new fit model (fit2) for h=30 periods (30 days).

```{r}
fcast <- forecast(fit2, h=30)
plot(fcast)
```

<a id="test-model-performance"></a>
### Test model performance with a holdout set

```{r}
hold <- window(ts(deseasonal_cnt), start(350))
fit_no_holdout = arima(deseasonal_cnt[-c(350:384)], order=c(1,1,7))
fcast_no_holdout <- forecast(fit_no_holdout, h=35)
plot(fcast_no_holdout, main="")
lines(ts(deseasonal_cnt))
```

Model needs seasonality added back in as it is too linear and not reality conform

```{r}
fit_w_seasonality <- auto.arima(deseasonal_cnt, seasonal = TRUE)
seas_forecast <- forecast(fit_w_seasonality, h=50)
plot(seas_forecast)

lines(ts(count_ma))
lines(ts(deseasonal_cnt))
```

<a id="further-analysis"></a>

### Futher analysis and testing of our models

Test against original auto.arima() p, d, q values

```{r}
fit3 <- auto.arima(deseasonal_cnt, seasonal=FALSE)
tsdisplay(residuals(fit3), lag.max=15, main ="Seasonal Model Residuals")
```

Evidence exists of a lag at 7, a higher order q value (q=7) might be necessary.

```{r}
fit4 <- arima(deseasonal_cnt, order = c(1,1,7)) #(2,1,7)
tsdisplay(residuals(fit4), lag.max=15, main ="Seasonal Model Residuals")

```

Default ARIMA model 1,1,1 (p,d,q) values for most apps

```{r}
fit5 <- arima(deseasonal_cnt, order=c(1,1,1))
tsdisplay(residuals(fit5), lag.max=15, main = "Seasonal Model Residuals")
```

Final fit and testing of the ARIMA forecast

```{r}
par(mfrow=c(2,2))

# auto arima (2,0,3) fit with seasonality
facst <- forecast(fit_w_seasonality, h=30)
plot(fcast)

# auto arima (2,0,3) fit without seasonality
fcast2 <- forecast(fit3, h=30)
plot(fcast2)

# custom arima (1,1,7) fit
fcast3 <- forecast(fit4, h=30)
plot(fcast3)

# general default arima (1,1,1) fit 
fcast4 <- forecast(fit5, h=30)
plot(fcast4)


```


<a id="sources"></a>

### Sources 
<a href="#top">back to top</a>

A Complete Tutorial on Time Series Modeling in R: https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/ <br>
ARIMA Forecasting in R Part 1 - Exploratory Data Analysis: https://www.youtube.com/watch?v=UekOBfpu8m8 <br>
ARIMA Forecasting in R Part 2 - Decomposition of the Data: https://www.youtube.com/watch?v=-qVAeU-_t_4 <br>
ARIMA Forecasting in R Part 3 - Autocorrelations and Choosing the Model Order: https://www.youtube.com/watch?v=qqAyMwYEIaA <br>
ARIMA Forecasting in R Part 4 - Fitting an ARIMA Model: https://www.youtube.com/watch?v=GpO89Hx088o <br>
ARIMA Forecasting in R Part 5 - Testing the ARIMA Model with a Holdout Set and Seasonality: 
https://www.youtube.com/watch?v=3I_NrpWsVWM <br>
ARIMA Forecasting in R Part 6 - Further Testing and Final Model Selection and Visualization: https://www.youtube.com/watch?v=k6aiuu9oiTY <br>
Applied Time Series Analysis for Fisheries and Environmental Sciences: https://nwfsc-timeseries.github.io/atsa-labs/ <br>
Introduction to Forecasting with ARIMA in R:   https://www.datascience.com/blog/introduction-to-forecasting-with-arima-in-r-learn-data-science-tutorials <br>